Cauchy example

### 2020-03-24 src/s_cauchy.Rmd
added cauchy convergence plots

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(mvtnorm)
library(goftest)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")
```

@@

## 10,000 iterations
Nominal model with 10000 iterations and thinning by 10
```{r}
fit <- stan(file = 'cauchy_nom.stan', seed = 7878,
            refresh = 0, iter = 1000)
full_data <- rstan::extract(fit, permuted=F)

@@

Alternative model with 10000 iterations and thinning by 10
```{r}
fit <- stan(file = 'cauchy_alt_1.stan', seed = 7878,
            refresh = 0, iter = 1000)
full_data <- rstan::extract(fit, permuted=F)

@@

>>>>

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(mvtnorm)
library(goftest)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
source("r_star_monitor.R")

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

@@

## 10,000 iterations
Nominal model with 10000 iterations and thinning by 10
```{r}
fit <- stan(file = 'cauchy_nom.stan', seed = 7878,
            refresh = 0, iter = 50000)
saveRDS(fit, "../output/fit_nominal_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

@@

Alternative model with 10000 iterations and thinning by 10
```{r}
fit <- stan(file = 'cauchy_alt_1.stan', seed = 7878,
            refresh = 0, iter = 50000)
saveRDS(fit, "../output/fit_alt_50000.rds")
full_data <- rstan::extract(fit, permuted=F)

@@

Compare quantiles
```{r}
fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
r_squared <- function(x, lower=0.001){
  qs <- seq(lower, 1-lower, lower)
  q_actual <- map_dbl(qs, ~qcauchy(.))
  r_vals <- vector(length = 50)
  for(i in seq_along(r_vals)){
    q_approx <- quantile(x[, , i], qs)
    aLM <- lm(data=tibble(q_approx, q_actual), formula = q_actual~q_approx)
    a_sum <- summary(aLM)
    r_vals[i] <- a_sum$r.squared
  }
  return(r_vals)
}

log_last <- log10(25000)
log_first <- log10(100)
iter <- round(10^(seq(log_first, log_last, length.out = 20)))
r_vals1 <- map_dbl(iter, ~median(r_squared(x[1:., , ])))
r_vals1_lower <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.025))
r_vals1_upper <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.975))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
r_vals <- map_dbl(iter, ~median(r_squared(x[1:., , ])))
r_vals_lower <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.025))
r_vals_upper <- map_dbl(iter, ~quantile(r_squared(x[1:., , ]), 0.975))
```

Compare nominal and alt parameterisations
```{r}
b_df <- 
  tibble(iteration=iter,
         middle=r_vals,
         lower=r_vals_lower,
         upper=r_vals_upper,
         variable="alt") %>% 
  bind_rows(tibble(iteration=iter,
         middle=r_vals1,
         lower=r_vals1_lower,
         upper=r_vals1_upper,
         variable="nom"))
# saveRDS(b_df, "../output/cauchy_r_squared.rds")
g1 <- b_df %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=middle,colour=as.factor(variable))) +
  geom_line(aes(y=middle,colour=as.factor(variable))) +
  ylim(NA, 1) +
  ylab(TeX("$R^2$")) +
  scale_color_brewer("Parameterisation", palette = "Dark2", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=11, colour="black"),
        axis.text = element_text(size=10, colour="black"),
        legend.position = "none",
        axis.title.y = element_blank()) +
  scale_x_log10() +
  ggtitle(TeX("Quantile $R^2$"))
g1
```

Progression of R*
```{r}
fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")

library(furrr)
future::plan(multiprocess)
laccuracy <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")

laccuracy1 <- future_map_dbl(iter, function(i) r_star(x[1:i, , ]))

a_df <- 
  tibble(iteration=iter,
         alt=laccuracy1,
       nom=laccuracy)
# saveRDS(a_df, "../output/cauchy_r_star.rds")
```

```{r}
g2 <- 
  a_df %>% 
  melt(id.vars="iteration") %>% 
  ggplot(aes(x=iteration, y=value, colour=as.factor(variable))) +
  geom_point() +
  geom_line() +
  ylab(TeX("$R^*$")) +
  scale_color_brewer("Parameterisation", palette = "Dark2", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=11, colour="black"),
        axis.text = element_text(size=10, colour="black"),
        legend.position = c(0.8, 0.8),
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_x_log10() +
  scale_y_continuous(limits=c(0.8, NA), breaks = seq(1, 7, 1)) +
  ggtitle(TeX("Split-$R^*$"))
g2
```

Rhat
```{r}
r_hat_all_params <- function(x){
 nparams <- dim(x)[3] 
 rhat_vals <- map_dbl(seq(1, nparams, 1), ~rhat(split_chains(x[ , , .])))
 return(rhat_vals)
}

fit <- readRDS("../output/fit_nominal_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
future::plan(multiprocess)
r_hat <- future_map_dbl(iter, ~max(r_hat_all_params(x[1:., , ])))
r_hat_lower <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.025))
r_hat_upper <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.975))

fit1 <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit1, permuted=F, pars="x")
r_hat1 <- future_map_dbl(iter, ~max(r_hat_all_params(x[1:., , ])))
r_hat1_lower <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.025))
r_hat1_upper <- future_map_dbl(iter, ~quantile(r_hat_all_params(x[1:., , ]), 0.975))

c_df <- 
  tibble(iteration=iter,
         middle=r_hat1,
         lower=r_hat1_lower,
         upper=r_hat1_upper,
         variable="alt") %>% 
  bind_rows(tibble(iteration=iter,
         middle=r_hat,
         lower=r_hat_lower,
         upper=r_hat_upper,
         variable="nom"))
# saveRDS(c_df, "../output/cauchy_r_hat.rds")
g3 <- 
  c_df %>% 
  ggplot(aes(x=iteration)) +
  geom_point(aes(y=middle, colour=as.factor(variable))) +
  geom_line(aes(y=middle, colour=as.factor(variable))) +
  ylab(TeX("$\\hat{R}$")) +
  scale_color_brewer("Parameterisation", palette = "Dark2", labels=c("Alternative", "Nominal")) +
  scale_fill_brewer("Parameterisation", palette = "Dark2", labels=c("Alternative", "Nominal")) +
  xlab("Iteration") +
  theme(text=element_text(size=11, colour="black"),
        axis.text = element_text(size=10, colour="black"),
        legend.position = "none",
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_x_log10() +
  scale_y_continuous(limits=c(0.99, NA), breaks = seq(1, 6, 0.2)) +
  ggtitle(TeX("Max split-$\\hat{R}$"))
g3
```

Combine
```{r}
pdf("../output/cauchy_convergence.pdf", width = 12, height = 4)
multiplot(g1, g3, g2, cols = 3)
dev.off()
```

Look at quantile accuracy for models converged by R* and Rhat
```{r}
fit <- readRDS("../output/fit_alt_50000.rds")
x <- rstan::extract(fit, permuted=F, pars="x")
quantile_errors <- function(x, lower=0.001){
  print(dim(x))
  qs <- seq(lower, 1-lower, lower)
  q_actual <- map_dbl(qs, ~qcauchy(.))
  errors <- matrix(nrow = length(qs),
                   ncol=dim(x)[3])
  for(i in 1:50){
    q_approx <- quantile(x[, , i], qs)
    errors[, i] <- abs(q_actual - q_approx)
  }
  errors <- errors %>% 
    as.data.frame() %>% 
    mutate(q=0.5-abs(0.5-qs)) %>% 
    melt(id.vars="q")
  return(errors)
}

errors_df <- quantile_errors(x, lower=0.001)
errors_df %>% 
  ggplot(aes(x=q, y=value)) +
  geom_jitter(width = 0.05) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method="lm") +
  xlab("Distance from centre") +
  ylab("Error")

aLM <- lm(data=errors_df, formula="log(value)~log(q)")
summary(aLM)

errors_df <- quantile_errors(x[1:iter[8], , ], lower=0.001)
errors_df %>% 
  ggplot(aes(x=q, y=value)) +
  geom_jitter(width = 0.05) +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method="lm") +
  xlab("Distance from centre") +
  ylab("Error")

bLM <- lm(data=errors_df, formula="log(value)~log(q)")
summary(bLM)
```
